\svnidlong
{$HeadURL$}
{$LastChangedDate$}
{$LastChangedRevision$}
{$LastChangedBy$}
\svnid{$Id$}

Dark matter\sidenote{Many theories of physics beyond the Standard Model predict the existence
of stable, neutral, weakly-interacting and massive particles that are
putative dark matter candidates. In the following, we refer to such
matter as dark matter, even though the observation of such matter at a collider
could only establish that it is neutral, weakly-interactive, massive and stable
on the distance-scales of tens of meters.} has not yet been observed in particle physics experiments, and
there is not yet any evidence for non-gravitational interactions
between dark matter and Standard Model particles.  If such
interactions exist, particles of dark matter could be produced
at the LHC. Since dark matter particles themselves do not produce signals
in the LHC detectors, one way to observe them is when they are produced in association
with a visible SM particle X(=$g, q, \gamma, Z, W$, or $h$).
%% so simple pair-production of dark matter particles, for example,
%% would
%% not be visible.
%% However, there are different mechanisms by which a
%% could be produced in association with dark matter.
Such reactions, which are
observed at colliders as particles or jets recoiling against an invisible state, are
called ``mono-X'' or \MET{}+X reactions~\cite{Feng:2005gj,Beltran:2010ww,Bai:2010hh}, 
where \MET is the missing transverse momentum observable in the detector.



Early Tevatron and LHC Run-1 searches for \MET{}+X signatures at CDF
CDF~\cite{Aaltonen:2012jb}, 
ATLAS~\cite{Aad:2015zva,Aad:2014tda,ATLAS:2014wra,Aad:2014vka,Aad:2013oja,Aad:2014wza,Aad:2014vea,ATL-PHYS-PUB-2014-007}
and
CMS~\cite{Khachatryan:2014rra,Khachatryan:2014rwa,Khachatryan:2014tva,Khachatryan:2014uma,Khachatryan:2015nua,CMS-PAS-B2G-13-004,CMS-PAS-EXO-14-004},
employed a basis of contact interaction operators in effective field
theories (EFTs) \cite{Goodman:2010yf,Goodman:2010ku} to calculate the
possible signals. %\Todo{Check these references here}
These EFTs assume that production of dark matter takes place through a
contact interaction involving a quark-antiquark pair, or two gluons,
and two dark matter particles.  In this case, the missing energy
distribution of the signal is determined by the nature and the mass of
the dark matter particles and the Lorentz structure of the
interaction. Only the overall production rate is a free parameter to
be constrained or measured.  Provided that the contact interaction
approximation holds, these EFTs provide a straightforward way to
compare the results from different collider searches with non-collider
searches for dark matter.  

The EFT describes the case when the mediator of the interaction between SM and DM particles are very heavy; 
if this is not the case, models that explicitly include these mediators are needed~\cite{Goodman:2010yf,Shoemaker:2011vi,Bai:2010hh,Kopp:2011eu,Fox:2011fx,Fox:2011pm,Shoemaker:2011vi,Busoni:2013lha}.
Some ``simplified models'' \cite{Alwall:2008ag,Goodman:2011jq,Alves:2011wf}
of dark matter production were constructed, including particles and interactions beyond the SM.
These models can be used consistently at LHC energies, and provide
an extension to the EFT approach. 
Many proposals for such models have emerged (see, for example
Refs. \cite{An:2012va,An:2012ue,Tait:2013,Buchmueller:2013dya,Bai:2013iqa,Bai:2014osa,An:2013xka,Yavin:14092893,Malik:2014ggr,Harris:2014hga,Buckley:2014fba,Haisch:2015ioa,Bai:2012xg,Carpenter:2012rg,Bell:2012rg,Petrov:2013nia,Carpenter:2013xra}). 
%% that a contact
%% interaction is often not the correct description for the signals to
%% which the LHC is sensitive.
%Limits from the LHC on particles that are
%QCD singlets and do not decay to hard leptons are relatively
%weak (see, for example, limits on dijet resonances for weakly-coupled
%theories).
%Such particles can mediate reactions that produce dark matter
%particles.
At the LHC, the kinematics of mono-X reactions occurring via a \tev-scale mediator can differ substantially from the prediction of the contact
interaction. The mediator may also produce qualitatively different signals, such as decays back into Standard Model particles. 
Thus, appropriate simplified models are thus an important component of the design, optimization, and interpretation of dark matter searches at ATLAS and CMS.
This has already been recognized in the CDF, ATLAS and CMS searches quoted above, where both EFT and selected simplified model
results are presented. 

\section{The ATLAS/CMS Dark Matter Forum}

To understand what signal models should be considered for the upcoming LHC Run 2, groups of experimenters from 
both ATLAS and CMS collaborations have held separate discussions with small groups of
theorists~\cite{Malik:2014ggr,Yavin:14092893}. 
Another joint discussion was held at the DM@LHC workshop in Oxford, in 
September 2014~\cite{DMatLHCProceedings}. 
Following the DM@LHC workshop, ATLAS and CMS organized a forum, called the~\textit{ATLAS-CMS Dark
Matter Forum}, to form a consensus on the use of simplified models
and EFTs for Run-2 with the participation of experts on
theories of dark matter. This is the final report of that forum.

One of the guiding principles of this report is to channel the efforts
of the ATLAS and CMS collaborations towards a minimal basis of dark
matter models that should influence the design of the early Run-2
searches. At the same time, a thorough survey of realistic collider
signals of dark matter is a crucial input to the overall design of the
search program.

The goal of this report is such a survey, though confined within some
broad assumptions and focused on benchmarks for kinematically-distinct
signals which are most urgently needed. As far as time and resources
have allowed, the assumptions have been carefully motivated by
theoretical consensus and comparisons of simulations. But, to achieve a
true consensus in only a few months before the start of Run-2, it was
important to restrict the scope and timescale to the following:

\begin{enumerate}
\item The forum should propose a prioritized, compact set of benchmark
  simplified models that should be agreed upon by both collaborations for
  Run-2 searches. The values for the scan on the parameters of the models for which
  experimental results are provided should be specified, to facilitate theory reinterpretation 
  beyond the necessary model-independent limits that 
  should be provided by all LHC Dark Matter searches. 
\item The forum should standardize the event generator implementation
  of the simplified models and harmonize other common technical
  details as far as practical. It
  would be desirable to have a common choice of leading order (LO) and 
  next-to-leading order (NLO) matrix elements corresponding to the state of the art calculations, 
  parton shower matching and merging, factorization and renormalization
  scales for each of the simplified models. This will also lead to a
  common set of theory uncertainties, which will facilitate the
  comparison of results between the two collaborations.
\item The forum should discuss how to apply the
  EFT formalism and present the results of EFT
  interpretations.
\item The forum should prepare a report summarizing these items,
  suitable both as a reference for the internal ATLAS and CMS
  audiences and as an explanation of Run-2 benchmarks for theory and non-collider
  readers. This report represents the views of its endorsers, as participants of the forum.
\end{enumerate}

\section{Grounding Assumptions}

We assume that interactions exist between Standard Model hadrons
and the particles that constitute cosmological dark matter. If this
is not the case, then proton collisions will not produce dark matter
particles, and dark matter will not scatter off nuclei in direct
detection experiments.

The Dark Matter itself is assumed to be a single particle, a Dirac
fermion WIMP, stable on collider timescales and non-interacting with
the detector.  
The former assumption is reductionistic.
The rich particle content of the Standard Model is circumstantial evidence that
the dark matter sector, which constitutes five times as much of the
mass of the universe, may be more complex than a single particle or a
single interaction. But, as was often the case in the discoveries of
the SM, here only one mediator and one search channel might play a
dominant role in the opening stages of an LHC discovery. The latter
assumption focuses our work on early LHC searches, where small
kinematic differences between models will not matter in a discovery
scenario, and with the imminent re-start of the LHC our report relies
heavily on a large body of existing theoretical work which assumed Dirac fermionic dark matter. 

Different spins of dark matter particles will typically
give similar results. Some exceptions exist. For example, the choice of Majorana fermions forbids some
processes that are allowed for Dirac fermions~\cite{Goodman:2010yf}.
Aside from these, adjusting the choice of Dirac or Majorana fermions or scalars will produce only minor changes
in the kinematic distributions of the visible particle and is expected to have little effect
on cut-and-count\sidenote{Cut-and-count refers to an analysis
that applies a certain event selection and checks the inclusive number of events which pass. 
This is to be contrasted with a shape analysis, which compares the distribution of events.} analysis. Thus the choice of Dirac
fermion dark matter should be sufficient as benchmarks for the upcoming Run-2 searches. 

% Should LHC searches uncover a signal consistent with these models, a much more exhaustive Nevertheless, a more complete
% set of models will certainly be required upon a discovery. See
% Refs.~\cite{Cotta:2012nj,Haisch:2013fla,Crivellin:2015wva} for some
% studies of observables that may distinguish amongst these models.

One advantage of collider experiments lies in their ability to study
and possibly characterize the mediator. A discovery of an anomalous
\MET signature at the LHC would not uniquely imply discovery of dark
matter, while at the same time discovery of an anomalous and
annually-modulated signal in a direct-detection experiment would leave
unanswered many questions about the nature of the interaction that
could be resolved by the simultaneous discovery of a new mediator
particle.  Collider, direct, and indirect detection searches provide
complementary ways to approach this problem~\cite{Bauer:2013ihz}, and it is in this spirit
that much of our focus is on the mediator.

We systematically explore the basic possibilities for
mediators of various possible spins and couplings.
All models considered are assumed to produce a signature with pairs of dark matter particles.
Though more varied and
interesting possibilities are added to the literature almost daily,
these basic building blocks account for much of the physics studied at
hadron colliders in the past three decades.

We also assume that Minimal Flavor Violation (MFV) \cite{Chivukula:1987py,Hall:1990ac,Buras:2000dm,D'Ambrosio:2002ex} applies to the
models included in this report. This means that the flavor structure of the
couplings between dark matter and ordinary particles follows the same
structure as the Standard Model. This choice is simple, since no
additional theory of flavor is required, beyond what is already
present in the SM, and it provides a mechanism to ensure that the
models do not violate flavor constraints.  As a consequence, \spinzero
resonances must have couplings to fermions proportional to the SM Higgs couplings. 
Flavor-safe models can still be constructed beyond the MFV
assumption, for example ~\cite{Agrawal:2014aoa}, and deserve further study.
%cite Dolan:2014ska? Also DM@LHC report, if out in time. 

In the parameter scan for the models considered in this report, we make the
assumption of a minimal decay width for the particles mediating the
interaction between SM and DM.  This means that only decays
strictly necessary for the self-consistency of the model (e.g.  to DM
and to quarks) are accounted for in the definition of the mediator
width. We forbid any further decays to other invisible particles of
the Dark Sector that may increase the width or produce striking, visible signatures. 
Studies within the Forum
show that, for cut-and-count analyses, the kinematic distributions of
many models, and therefore the sensitivity of these searches, do not depend
significantly on the mediator width, as long as the width remains smaller
than the mass of the particle and that narrow mediators are sufficiently light.

The particle content of the models chosen as benchmarks is limited to
one single kind of DM whose self-interactions are not relevant for LHC
phenomenology, and to one type of SM/DM interaction at a time. These
assumptions only add a limited number of new particles to the
SM. These simplified models, independently explored by different
experimental analyses, can be used as starting points to build more
complete theories. Even though this factorized picture does not always
lead to full theories and leaves out details that are necessary for
the self-consistency of single models (e.g. the mass generation for
mediator particles), it is a starting point to prepare a set of
distinct but complementary collider searches for Dark Matter, as
it leads to benchmarks that are easily comparable across channels.

\section{Choices of benchmarks considered in this report and parameter scan}

%We have already said this above. 
%carts in front of the horse?
%The weakness of the interaction between dark matter and visible matter is explained
%by assuming a mediating force or particle that indirectly connects the
%two. In the EFT approach, the mass scale of the mediator is assumed
%to be large compared to the typical momentum involved in the production
%of dark matter.  
Contact interaction operators have been outlined as basis set of theoretical
building blocks representing possible types of interactions between SM and DM particles
in~\cite{Goodman:2010ku}. The approach followed by LHC searches (see e.g. Refs.~\cite{Khachatryan:2014rra,Aad:2015zva} 
for the most recent jet+\MET{} Run-1 searches with the 8 TeV dataset) 
so far has been to simulate only a prioritized set of the possible operators with distinct kinematics
for the interpretation of the constraints obtained, and provide results that may be reinterpreted in terms of the other operators.
This report intends to follow this strategy, firstly focusing on simplified models that allow the exploration 
of scenarios where the mediating scale is not as large.  In the limit of large mediator mass, the simplified models map onto
the EFT operators.
%because the EFT approach, viable for the
%other types of dark matter searches, breaks down at the LHC. Another
%way of viewing this problem is that any BSM model of dark matter
%requires at least two particles: a dark matter candidate and a
%mediator of the interaction between the dark matter and the Standard
%Model.
Secondly, this report considers specific EFT benchmarks 
whenever neither a simplified model completion 
nor other simplified models yielding similar kinematic distributions are available\sidenote{and implemented in one of the event generators used by both collaborations.}. 
This is the case for dimension-5 or dimension-7\sidenote{For dim-7 operators of 
DM coupling to gauge bosons, there are simplified models in the literature, but they need more ingredients than just the usual mediator and the DM. For example, consider the DM coupling to a SM Higgs singlet mediator, $S$, whose vaccuum expectatino value gives mass to a heavy vector-like fermion, $F$, which also carries SM quantum numbers. The $F$ triangle graph then couples $S$ to the SM gauge boson pair. Thus $\chi \bar{\chi} \rightarrow S \rightarrow F\bar{F} \text{(loop)} \rightarrow G G$ as required.} operators with direct 
Dark Matter-electroweak boson couplings. 
Considering these models as separate experimental benchmarks 
will allow to target new signal regions and help validate the 
contact interaction limit of new simplified models 
developed to complete these specific operators. 

In all cases, experimental searches should deliver 
results that are independent from the specific benchmark tested. 
Furthermore, it should be possible to map results from simplified models to contact operators,
in the high mediator mass limit. Results from EFT benchmarks should include the condition that
the momentum transfer does not probe the scale of the interaction; whenever there is no model
that allows a direct mapping between these two quantities, various options should be tested to 
ensure a given fraction of events within the range of applicability of the EFT approach.

This report does not quantitatively consider constraints 
'external' to the MET+X analyses coming from
the LHC experiments themselves, such as searches for mediator decays,
when choosing the points to be scanned in the parameter space of the models. 
The main reason for not doing so in this report 
is the difficulty of incorporating these constraints in a rigorous quantitative way within
the timescale of the Forum. However, even if the parameter scans
and the searches are not optimized with those constraints in mind, 
we intend to make all information available to the community to exploit
the unique sensitivity of colliders to all possible Dark Matter signatures. 

\section{Structure of this report}

Chapter \ref{subsec:MonojetLikeModels} of the report is dedicated to simplified
models with radiation of a hard object either from the initial state
or from the mediator. These models produce primarily monojet signatures, 
but should be considered for all \MET{}+X searches.
%Chapter~\ref{subsec:DMHFModels} describes the specific details of the general models
%for final states including heavy flavor quarks, leaving the implementation
%details to appendix~\ref{app:TTBar_Xsecs}.  
Chapter~\ref{subsec:EWSpecificModels} contains the benchmark model
recommendations and choices for final states specifically containing an electroweak 
boson ($W/Z/\gamma/H$). In this case, both 
simplified models leading to mono-boson signatures
and contact interaction operators are considered. 
Details of the implementation of the simplified models in
Monte Carlo generators are provided in
Chapter \ref{app:MonojetLikeModels_Appendix}.
Chapter \ref{sec:EFTValidity} is devoted to the treatment of the presentation of results for the benchmark
models from contact interaction operators. Chapter \ref{sec:TheoryUncertainties} prescribes how to estimate theoretical uncertainties on the simulation of these models. Chapter \ref{chapter:conclusions} concludes the report.

The technical implementation and further models that can be studied
beyond early searches are described in Appendix~\ref{app:EWSpecificModels_Appendix}. 
For these models, either the implementation could not be fully developed by the time of this report,
or some of the grounding assumptions were not fully met.  
Some of these models have been used in previous ATLAS and CMS analyses and discussed thoroughly within the Forum. 
They are therefore worth considering for further studies and for Run-2 searches, since they lead to unique \MET{}+X signatures 
that are not produced by any other of the models included in this report. 

Appendix~\ref{app:Presentation_Of_Experimental_Results} contains the necessary elements that
should be included in the results of experimental searches to allow for further reinterpretation. 
%The Appendices also collect information about the implementation of the
%models in event generator tools, as well as tables of expected cross sections.
%%This could be saved for monotop

